---
title: "MARS and Random Forest"
author: "Sarah AlQahtani"
date: "11/17/2020"
output: html_document
---

```{r setup, include=FALSE}
# Set global R options
options(scipen = 999)

# Set the graphical theme
ggplot2::theme_set(ggplot2::theme_light())

# Set global knitr chunk options
knitr::opts_chunk$set(
  cache = TRUE,
  warning = FALSE, 
  message = FALSE
)
```
Using the same dataset leveraged for Portfolio Builder Exercise #1 & #2, write up a third report that answers the following:
```{r}
# Helper packages
library(dplyr)     # for data wrangling
library(ggplot2)   # for awesome plotting

# Modeling packages
library(earth)     # for fitting MARS models
library(caret)     # for automating the tuning process
library(ranger)   # a c++ implementation of random forest 
library(h2o)      # a java-based implementation of random forest
library(rsample)  # resampling procedures

# Model interpretability packages
library(vip)       # for variable importance
library(pdp)       # for variable relationships


```

1. Apply a MARS model with all features.
```{r}
# Fit a basic MARS model
mars <- earth(
  total_UPDRS ~ .,  
  data = par_train   
)

# Print model summary
print(mars)
```
```{r}
summary(mars) %>% .$coefficients %>% head(10)
```
```{r}
plot(mars, which = 1)
```
## Tuningg
```{r}
seq(2, 100, length.out = 10) %>% floor()
```

```{r}
# create a tuning grid
hyper_grid <- expand.grid(
  degree = 1:3, 
  nprune = seq(2, 100, length.out = 10) %>% floor()
)

hyper_grid
```
```{r}
# Cross-validated model
set.seed(123)  # for reproducibility
cv_mars <- train(
  x = subset(par_train, select = -total_UPDRS),
  y = par_train$total_UPDRS,
  method = "earth",
  metric = "RMSE",
  trControl = trainControl(method = "cv", number = 10),
  tuneGrid = hyper_grid
)

# View results
cv_mars$bestTune
```
```{r}
cv_mars$results %>%
  filter(nprune == cv_mars$bestTune$nprune, degree == cv_mars$bestTune$degree)
```
The cross-validated RMSE for these models is displayed below; the optimal model’s cross-validated RMSE is 1.633541.
```{r}
ggplot(cv_mars)
```


- How does the model performance compare to your previous models?
```{r}
# extract out of sample performance measures
summary(resamples(list(
  "LM" = cv_model2, 
  "PCR" = cv_model_pcr, 
  "PLS" = cv_model_pls,
  "Regularized Regression " = cv_glmnet,
  "MARS" = cv_mars
)))$statistics$RMSE %>%
  kableExtra::kable(caption = "Cross-validated RMSE results for tuned MARS and regression models.") %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"))
```
The MARS model higher  RMSE than the regression models but have a higher RMSE.

- How many of the features are influential? Which 10 features are considered most influential?
```{r}
# variable importance plots
p1 <- vip(cv_mars, num_features = 10, geom = "point", value = "gcv") + ggtitle("GCV")
p2 <- vip(cv_mars, num_features = 10, geom = "point", value = "rss") + ggtitle("RSS")

gridExtra::grid.arrange(p1, p2, ncol = 2)
```
as we can see above the following features are the most influential features:
  - motor_UPDRS
  - subject
  - age
  - sex
  - HNR
  - Jitter.Abs.
  - DFA
- Does your model include hinge functions? If so, explain their coefficient and plot their impact on the predicted response variable.
```{r}
summary(cv_mars) %>% .$coefficients %>% head(10)
```
- Does your model include interactions? If so, pick the interaction effect that is most influential and explain the coefficient.
```{r}
# extract coefficients, convert to tidy data frame
cv_mars$finalModel %>%
  coef() %>%  
  broom::tidy()
```
As our model perform the best when degree=1, we can see that that there are no interaction terms.

2. Apply a random forest model.
- First, apply a default random forest model.
```{r}
# number of features
n_features <- length(setdiff(names(par_train), "total_UPDRS"))

# train a default random forest model
par_rf1 <- ranger(
  total_UPDRS ~ ., 
  data = par_train,
  mtry = floor(n_features / 3),
  respect.unordered.factors = "order",
  seed = 123
)

# get OOB RMSE
(default_rmse <- sqrt(par_rf1$prediction.error))
```

- Now apply a a full cartesian grid search across various values of mtry, tree complexity & sampling scheme.
```{r}
# create hyperparameter grid
hyper_grid <- expand.grid(
  mtry = floor(n_features * c(.05, .15, .25, .333, .4)),
  min.node.size = c(1, 3, 5, 10), 
  replace = c(TRUE, FALSE),                               
  sample.fraction = c(.5, .63, .8),                       
  rmse = NA                                               
)

# execute full cartesian grid search
for(i in seq_len(nrow(hyper_grid))) {
  # fit model for ith hyperparameter combination
  fit <- ranger(
    formula         = total_UPDRS ~ ., 
    data            = par_train, 
    num.trees       = n_features * 10,
    mtry            = hyper_grid$mtry[i],
    min.node.size   = hyper_grid$min.node.size[i],
    replace         = hyper_grid$replace[i],
    sample.fraction = hyper_grid$sample.fraction[i],
    verbose         = FALSE,
    seed            = 123,
    respect.unordered.factors = 'order',
  )
  # export OOB error 
  hyper_grid$rmse[i] <- sqrt(fit$prediction.error)
}

# assess top 10 models
hyper_grid %>%
  arrange(rmse) %>%
  mutate(perc_gain = (default_rmse - rmse) / default_rmse * 100) %>%
  head(10)
```

- Now run a random grid search across the same hyperparameter grid but restrict the time or number of models to run to 50% of the models ran in the full cartesian.
```{r}
h2o.no_progress()
h2o.init(max_mem_size = "5g")

```

```{r}
# convert training data to h2o object
train_h2o <- as.h2o(par_train)

# set the response column to Sale_Price
response <- "total_UPDRS"

# set the predictor names
predictors <- setdiff(colnames(par_train), response)
```

```{r}
h2o_rf1 <- h2o.randomForest(
    x = predictors, 
    y = response,
    training_frame = train_h2o, 
    ntrees = n_features * 10,
    seed = 123
)

h2o_rf1
```
```{r}
# hyperparameter grid
hyper_grid <- list(
  mtries = floor(n_features * c(.05, .15, .25, .333, .4)),
  min_rows = c(1, 3, 5, 10),
  max_depth = c(10, 20, 30),
  sample_rate = c(.55, .632, .70, .80)
)

# random grid search strategy
search_criteria <- list(
  strategy = "RandomDiscrete",
  stopping_metric = "mse",
  stopping_tolerance = 0.001,   # stop if improvement is < 0.1%
  stopping_rounds = 10,         # over the last 10 models
  max_runtime_secs = 60*5      # or stop search after 5 min.
)
```

```{r}
# perform grid search 
random_grid <- h2o.grid(
  algorithm = "randomForest",
  grid_id = "rf_random_grid",
  x = predictors, 
  y = response, 
  training_frame = train_h2o,
  hyper_params = hyper_grid,
  ntrees = n_features * 10,
  seed = 123,
  stopping_metric = "RMSE",   
  stopping_rounds = 10,           # stop if last 10 trees added 
  stopping_tolerance = 0.005,     # don't improve RMSE by 0.5%
  search_criteria = search_criteria
)

```

```{r}
# collect the results and sort by our model performance metric 
# of choice
random_grid_perf <- h2o.getGrid(
  grid_id = "rf_random_grid", 
  sort_by = "rmse", 
  decreasing = FALSE
)
random_grid_perf
```

3. Pick the best performing model from above.
Our grid search assessed __240__ models before stopping due to time. The best model (`max_depth = 30`, `min_rows = 1`, `mtries = 8`, and `sample_rate = 0.8`) achieved an OOB RMSE of 0.44707568204518283.
- Identify the most influential features for this model.
```{r}
# re-run model with impurity-based variable importance
rf_impurity <- ranger(
  formula = total_UPDRS ~ ., 
  data = par_train, 
  num.trees = 3000,
  mtry = 8,
  min.node.size = 1,
  sample.fraction = .80,
  replace = FALSE,
  importance = "impurity",
  respect.unordered.factors = "order",
  verbose = FALSE,
  seed  = 123
)

# re-run model with permutation-based variable importance
rf_permutation <- ranger(
  formula = total_UPDRS ~ ., 
  data = par_train, 
  num.trees = 3000,
  mtry = 8,
  min.node.size = 1,
  sample.fraction = .80,
  replace = FALSE,
  importance = "permutation",
  respect.unordered.factors = "order",
  verbose = FALSE,
  seed  = 123
)
```

- Plot the top 10 most influential features.
```{r}
p1 <- vip::vip(rf_impurity, num_features = 10, geom = "point")
p2 <- vip::vip(rf_permutation, num_features = 10, geom = "point")

gridExtra::grid.arrange(p1, p2, nrow = 1)
```

- Do these features have positive or negative impacts on your response variable?
```{r}

```

- Create partial dependence plots for these features. Explain the relationship between the feature and the predicted values.
```{r}
p1 <- partial(rf_impurity, pred.var = "motor_UPDRS", grid.resolution = 10) %>% 
  ggplot(aes(motor_UPDRS, yhat)) +
  geom_line()

p2 <- partial(rf_impurity, pred.var = "subject.", grid.resolution = 10) %>% 
  ggplot(aes(subject., yhat)) +
  geom_line()

p3 <- partial(rf_impurity, pred.var ="age", 
              grid.resolution = 10) %>% 
  ggplot(aes(age, yhat)) +
  geom_line()

# Display plots side by side
gridExtra::grid.arrange(p1, p2, p3, ncol = 3)
```

