---
title: "05-GBMs_AutoML"
author: "Sarah AlQahtani"
date: "11/21/2020"
output: html_document
---
```{r}
# Helper packages
library(dplyr)    # for general data wrangling needs

# Modeling packages
library(gbm)      # for original implementation of regular and stochastic GBMs
library(h2o)      # for a java-based implementation of GBM variants
library(xgboost)  # for fitting extreme gradient boosting
library(rsample)  # sampling procedure

```
Using the same dataset leveraged for Portfolio Builder Exercises #1, #2, & #3 write up a third report that answers the following:
1. Apply a basic GBM model with the same features you used in the random forest module.
- Apply the default hyperparameter settings with a learning rate set to 0.10. How does model performance compare to the random forest module?
```{r}
# run a basic GBM model
set.seed(123)  # for reproducibility
par_gbm1 <- gbm(
  formula = total_UPDRS ~ .,
  data = par_train,
  distribution = "gaussian",  # SSE loss function
  n.trees = 3000,
  shrinkage = 0.1,
  interaction.depth = 3,
  n.minobsinnode = 1,
  cv.folds = 10
)

# find index for number trees with minimum CV error
best <- which.min(par_gbm1$cv.error)

# get MSE and compute RMSE
sqrt(par_gbm1$cv.error[best])
```
```{r}
# plot error curve
gbm.perf(par_gbm1, method = "cv")

```

- How many trees were applied? Was this enough to stabilize the loss function or do you need to add more?
1564 trees. Yes it needs to be improved.

- Tune the hyperparameters using the suggested tuning strategy for basic GBMs. Did your model performance improve?
```{r}
hyper_grid <- expand.grid(
  learning_rate = c(0.3, 0.1, 0.05, 0.01, 0.005),
  RMSE = NA,
  trees = NA,
  time = NA
)

# execute grid search
for(i in seq_len(nrow(hyper_grid))) {

  # fit gbm
  set.seed(123)  # for reproducibility
  train_time <- system.time({
    m <- gbm(
      formula = total_UPDRS ~ .,
      data = par_train,
      distribution = "gaussian",
      n.trees = 3000, 
      shrinkage = hyper_grid$learning_rate[i], 
      interaction.depth = 3, 
      n.minobsinnode = 10,
      cv.folds = 10 
   )
  })
  
  # add SSE, trees, and training time to results
  hyper_grid$RMSE[i]  <- sqrt(min(m$cv.error))
  hyper_grid$trees[i] <- which.min(m$cv.error)
  hyper_grid$Time[i]  <- train_time[["elapsed"]]

}

# results
arrange(hyper_grid, RMSE)
```
After, we will set the learning rate to the optimal level which is 0.1 and tune the tree specific hyperparameters interaction.depth and n.minobsinnode.
```{r}
hyper_grid <- expand.grid(
  n.trees = 4000,
  shrinkage = 0.1,
  interaction.depth = c(3, 5, 7),
  n.minobsinnode = c(5, 10, 15)
)

# create model fit function
model_fit <- function(n.trees, shrinkage, interaction.depth, n.minobsinnode) {
  set.seed(123)
  m <- gbm(
    formula = total_UPDRS ~ .,
    data = par_train,
    distribution = "gaussian",
    n.trees = n.trees,
    shrinkage = shrinkage,
    interaction.depth = interaction.depth,
    n.minobsinnode = n.minobsinnode,
    cv.folds = 10
  )
  # compute RMSE
  sqrt(min(m$cv.error))
}

# perform search grid with functional programming
hyper_grid$rmse <- purrr::pmap_dbl(
  hyper_grid,
  ~ model_fit(
    n.trees = ..1,
    shrinkage = ..2,
    interaction.depth = ..3,
    n.minobsinnode = ..4
    )
)

# results
arrange(hyper_grid, rmse)

```

Now, we will take our best model hypermeters and reduce the number of the learnong rate to 0.001 and increase the number of trees 6000 to improve the accuracy.
```{r}
hyper_grid <- expand.grid(
  n.trees = 6000,
  shrinkage = 0.001,
  interaction.depth = c(3, 5, 7),
  n.minobsinnode = c(5, 10, 15)
)

# create model fit function
model_fit <- function(n.trees, shrinkage, interaction.depth, n.minobsinnode) {
  set.seed(123)
  m <- gbm(
    formula = total_UPDRS ~ .,
    data = par_train,
    distribution = "gaussian",
    n.trees = n.trees,
    shrinkage = shrinkage,
    interaction.depth = interaction.depth,
    n.minobsinnode = n.minobsinnode,
    cv.folds = 10
  )
  # compute RMSE
  sqrt(min(m$cv.error))
}

# perform search grid with functional programming
hyper_grid$rmse <- purrr::pmap_dbl(
  hyper_grid,
  ~ model_fit(
    n.trees = ..1,
    shrinkage = ..2,
    interaction.depth = ..3,
    n.minobsinnode = ..4
    )
)

# results
arrange(hyper_grid, rmse)

```
RMSE does not shoe ane improvement.


2. Apply a stochastic GBM model. Tune the hyperparameters using the suggested tuning strategy for stochastic GBMs. Did your model performance improve?
```{r}
# refined hyperparameter grid
hyper_grid <- list(
  sample_rate = c(0.5, 0.75, 1),              # row subsampling
  col_sample_rate = c(0.5, 0.75, 1),          # col subsampling for each split
  col_sample_rate_per_tree = c(0.5, 0.75, 1)  # col subsampling for each tree
)

# random grid search strategy
search_criteria <- list(
  strategy = "RandomDiscrete",
  stopping_metric = "mse",
  stopping_tolerance = 0.001,   
  stopping_rounds = 10,         
  max_runtime_secs = 60*60      
)

# perform grid search 
grid <- h2o.grid(
  algorithm = "gbm",
  grid_id = "gbm_grid",
  x = predictors, 
  y = response,
  training_frame = train_h2o,
  hyper_params = hyper_grid,
  ntrees = 4000,
  learn_rate = 0.01,
  max_depth = 7,
  min_rows = 10,
  nfolds = 10,
  stopping_rounds = 10,
  stopping_tolerance = 0,
  search_criteria = search_criteria,
  seed = 123
)

# collect the results and sort by our model performance metric of choice
grid_perf <- h2o.getGrid(
  grid_id = "gbm_grid", 
  sort_by = "mse", 
  decreasing = FALSE
)
```

```{r}
# Grab the model_id for the top model, chosen by cross validation error
best_model_id <- grid_perf@model_ids[[1]]
best_model <- h2o.getModel(best_model_id)

# Now let’s get performance metrics on the best model
h2o.performance(model = best_model, xval = TRUE)
```

3. Apply an XGBoost model. Tune the hyperparameters using the suggested tuning strategy for XGBoost models.
```{r}
library(recipes)
xgb_prep <- recipe(total_UPDRS ~ ., data = ames_train) %>%
  step_integer(all_nominal()) %>%
  prep(training = par_train, retain = TRUE) %>%
  juice()

X <- as.matrix(xgb_prep[setdiff(names(xgb_prep), "total_UPDRS")])
Y <- xgb_prep$total_UPDRS

```

```{r}
set.seed(123)
ames_xgb <- xgb.cv(
  data = X,
  label = Y,
  nrounds = 6000,
  objective = "reg:linear",
  early_stopping_rounds = 50, 
  nfold = 10,
  params = list(
    eta = 0.05,
    max_depth = 3,
    min_child_weight = 3,
    subsample = 0.8,
    colsample_bytree = 0.5),
  verbose = 0
)  

# minimum test CV RMSE
min(ames_xgb$evaluation_log$test_rmse_mean)
```

```{r}
# hyperparameter grid
hyper_grid <- expand.grid(
  eta = 0.05,
  max_depth = 3, 
  min_child_weight = 3,
  subsample = 0.8, 
  colsample_bytree = 0.5,
  gamma = c(0, 1, 10, 100, 1000),
  lambda = c(0, 1e-2, 0.1, 1, 100, 1000, 10000),
  alpha = c(0, 1e-2, 0.1, 1, 100, 1000, 10000),
  rmse = 0,          # a place to dump RMSE results
  trees = 0          # a place to dump required number of trees
)

# grid search
for(i in seq_len(nrow(hyper_grid))) {
  set.seed(123)
  m <- xgb.cv(
    data = X,
    label = Y,
    nrounds = 4000,
    objective = "reg:linear",
    early_stopping_rounds = 50, 
    nfold = 10,
    verbose = 0,
    params = list( 
      eta = hyper_grid$eta[i], 
      max_depth = hyper_grid$max_depth[i],
      min_child_weight = hyper_grid$min_child_weight[i],
      subsample = hyper_grid$subsample[i],
      colsample_bytree = hyper_grid$colsample_bytree[i],
      gamma = hyper_grid$gamma[i], 
      lambda = hyper_grid$lambda[i], 
      alpha = hyper_grid$alpha[i]
    ) 
  )
  hyper_grid$rmse[i] <- min(m$evaluation_log$test_rmse_mean)
  hyper_grid$trees[i] <- m$best_iteration
}
hyper_grid %>%
  arrange(rmse)
```

```{r}
# optimal parameter list
params <- list(
  eta = 0.05,
  max_depth = 3,
  min_child_weight = 3,
  subsample = 0.8,
  colsample_bytree = 0.5,
  lambda = 1,
  alpha = 10000
)

# train final model
xgb.fit.final <- xgboost(
  params = params,
  data = X,
  label = Y,
  nrounds = 1350,
  objective = "reg:linear",
  verbose = 0
)
```

- Did your model performance improve?
- Did regularization help?
4. Pick your best GBM model. Which 10 features are considered most influential? Are these the same features that have been influential in previous models?

5. Create partial dependence plots for the top two most influential features. Explain the relationship between the feature and the predicted values.

6. Using H2O, build and assess the following individual models:
- regularized regression base learner,
- random forest base learner.
- GBM and/or XGBoost base learner.
7. Using h2o.stackedEnsemble(), stack these three models.
- Does your stacked model performance improve over and above the individual learners?
- Explain your reasoning why or why not performance improves.
8. Perform a stacked grid search with an H2O GBM or XGBoost model.
- What was your best performing model?
- Do you notice any patterns in the hyperparameter settings for the top 5-10 models?
9. Perform an AutoML search across multiple types of learners.
- Which types of base learners are in the top 10?
- What model provides the optimal performance?
- Apply this model to the test set. How does the test loss function compare to the training cross-validated RMSE?
